---
title: "Testing Alternate Method to Compute Empirical p-Values"
author: "Kevin R. Coombes"
date: "26 April 2018"
output: 
  html_document:
    toc: true
    highlight: kate
    theme: yeti
---

```{r mycss, results="asis", echo=FALSE}
cat('
<style type="text/css">
b, strong {color: red; }
i, em {color: blue; }
.defn {color: purple; }
.para {color: purple;
      font-weight: bold;
}
.figure { text-align: center; }
.caption { font-weight: bold; }
</style>
')
```

# Introduction
This is the second profiling study. Even after improving (significantly) the
time taken by the 'ranNuGen' method, the code is still uncofortably slow. We
like to figure out where time is being taken.

This should only be run with a specific version of the package.
```{r vers}
library(NewmanOmics)
if (packageVersion("NewmanOmics") != "0.6.0") {
  stop("Wrong NewmanOmics package version")
}
```

# Read the Data
We use our luing cancer example for profiling.
```{r lungdata}
csvfile <- system.file("extdata", "LungNormalTumorPair.csv",
                       package="NewmanOmics")
lung <- read.csv(csvfile)
summary(lung)
```
# Profiling
Now we can profile the overall process.
```{r profile}
set.seed(12345)
Rprof("myProfile01.out")
for (ignore in 1:10) {
  picked <- sample(nrow(lung), 1000)
  normal <- log(1 + as.matrix(lung[picked, 2, drop=FALSE]))
  tumor  <- log(1 + as.matrix(lung[picked, 3, drop=FALSE]))
  ps <- pairedStat(normal, tumor)
  assign(paste("ps", ignore, sep='.'), ps, .GlobalEnv)
}
Rprof()
summaryRprof("myProfile01.out")
```
Basically, the time is now spent inside `mean` and `<`, which we can
locate inside the function `nu2PValPaired` function.  We should be able
to speed up the computation by using a binary search in place of the
existing construction, `mean(x < vec)`.

# Speeding Things Up
We use the following versionof a binary search tool. It was originally
written by Chao Liu on stackoverflow at
(https://stackoverflow.com/questions/20133344/find-closest-value-in-a-vector-with-binary-search)
We have modified it to 
```{r nvs}
NearestValueSearch <- function(x, w){
  ## A simple binary search algorithm
  ## Assume the w vector is sorted so we can use binary search
  left <- 1
  right <- length(w)
  while(right - left > 1){
    middle <- floor((left + right) / 2)
    if(x < w[middle]){
      right <- middle
    }
    else{
      left <- middle
    }
  }
  if(abs(x - w[right]) < abs(x - w[left])){
    return(right)
  }
  else{
    return(left)
  }
}
```

Here is how we incorporate that into the function that computes
empirical p-values.
```{r altalg}
altnu2PValPaired <- function(nuMatrix, vec){
  vec <- sort(vec)
  MatP <- matrix(sapply(nuMatrix, function(x) {
    1 - NearestValueSearch(x, vec)/length(vec)
  }), nrow(nuMatrix), ncol(nuMatrix))
  return(MatP)
}
```

Now let's see how long just this part takes with the new method.
```{r reprof}
set.seed(54321)
Rprof("myProfileAlt.out")
for (ignore in 1:10) {
  ps <- get(paste("ps", ignore, sep='.'), .GlobalEnv)
  numb <- NewmanOmics:::randNuGen()
  x2 <- altnu2PValPaired(ps$nu.statistics, numb)
  print(summary(ps$p.values - x2))
}
Rprof()
summaryRprof("myProfileAlt.out")
```

This is clearly **lots** faster.

# Appendix
```{r si}
sessionInfo()
```
