---
title: "Profiling the Banked Statistic"
author: "Kevin R. Coombes"
date: "26 April 2018"
output: 
  html_document:
    toc: true
    highlight: kate
    theme: yeti
---

```{r mycss, results="asis", echo=FALSE}
cat('
<style type="text/css">
b, strong {color: red; }
i, em {color: blue; }
.defn {color: purple; }
.para {color: purple;
      font-weight: bold;
}
.figure { text-align: center; }
.caption { font-weight: bold; }
</style>
')
```

# Introduction
This is the second profiling study. Even after improving (significantly) the
time taken by the 'ranNuGen' method, the code is still uncofortably slow. We
like to figure out where time is being taken.

This should only be run with a specific version of the package.
```{r vers}
library(NewmanOmics)
if (packageVersion("NewmanOmics") != "0.8.0") {
  stop("Wrong NewmanOmics package version")
}
```

# Read the Data
We use our head-and-neck cancer example for profiling.
```{r lungdata}
csvfile <- system.file("extdata", "GSE6631_Head&NeckSq.csv",
                       package="NewmanOmics")
HN <- as.matrix(read.csv(csvfile, row.names=1))
HN <- log2((1/19)*median(HN) + HN)
dim(HN)
colnames(HN)
```
# Profiling
Now we can profile the overall process.
```{r profile}
set.seed(12345)
Rprof("myProfile03.out")
for (ignore in 1:20) {
  picked <- sample(nrow(HN), 1500)
  bankData <- HN[picked, seq(1, ncol(HN), 2)]
  testSet <- HN[picked, seq(2, 6, 2)]
  bs <- bankStat(testSet = testSet, bankMatrix = bankData)
  assign(paste("bs", ignore, sep='.'), bs, .GlobalEnv)
}
Rprof()
summaryRprof("myProfile03.out")
```

# Appendix
```{r si}
sessionInfo()
```
